---
title: "Installation"
---

# Installation

didgpu is on CRAN and on r-universe. Pick whichever lane you prefer.

## From CRAN (recommended)

```{r install-cran, eval=FALSE}
install.packages("didgpu")
```

This gets you the pure-R and Rcpp backends. They are sufficient for
every estimator in this book; the chapters that follow do not require
GPU.

## From r-universe (newest)

If you want the latest commit on `main` before it lands on CRAN:

```{r install-runiverse, eval=FALSE}
install.packages(
  "didgpu",
  repos = c("https://joshuaammons.r-universe.dev", "https://cloud.r-project.org")
)
```

## From source (with GPU support)

If you have an NVIDIA GPU and the CUDA Toolkit installed, install from
source to compile the GPU kernels:

```{r install-source, eval=FALSE}
# (1) Set CUDA_HOME or CUDA_PATH to point at your CUDA install:
Sys.setenv(CUDA_HOME = "/usr/local/cuda")             # Linux
# or
Sys.setenv(CUDA_PATH = "C:/Program Files/NVIDIA GPU Computing Toolkit/CUDA/v13.2")  # Windows

# (2) Then install from source:
install.packages("didgpu", type = "source")           # CRAN source
# -- or --
remotes::install_github("JoshuaAmmons/didgpu")        # GitHub head
```

didgpu's configure script will print a one-line confirmation at install
time telling you whether it found CUDA:

```
*** didgpu configure ***
  CUDA toolkit detected: /usr/local/cuda
  -> Building WITH GPU support.
     backends available after install: 'r', 'cpu', 'cuda'
```

If CUDA is not found, the install still succeeds — only the `r` and
`cpu` backends are compiled and `backend = "cuda"` returns an
informative error at runtime. There is no CRAN-style ERROR or WARNING
during install in that case.

::: {.callout-note}
## "Did the GPU build actually work?"

After install, the fastest way to check is:

```{r check-cuda, eval=FALSE}
library(didgpu)
didgpu_has_cuda_support()
#> [1] TRUE
```

If you see `TRUE`, you can pass `backend = "cuda"` to any of the
estimator functions in this book. If you see `FALSE`, didgpu still
works in pure-R / Rcpp; only the GPU acceleration is unavailable.
:::

## Supported GPUs

didgpu compiles for compute capabilities **sm_75** through **sm_120**,
which covers:

| Architecture | Compute capability | Example cards |
|---|---|---|
| Turing | sm_75 | RTX 20-series, T4, GTX 16-series |
| Ampere (data center) | sm_80 | A100 |
| Ampere (desktop) | sm_86 | RTX 30-series |
| Ada Lovelace | sm_89 | RTX 40-series, RTX 4000 Ada |
| Hopper | sm_90 | H100 |
| Blackwell | sm_120 | RTX PRO Blackwell, RTX 50-series |

There is also a `compute_120` PTX target so newer Blackwell-and-after
cards JIT-compile from PTX cleanly.

## System requirements

* R ≥ 4.1
* On Linux/macOS: GNU make (already on most systems)
* On Windows: [Rtools45](https://cran.r-project.org/bin/windows/Rtools/)
  (only when building from source)
* **Optional** for GPU: NVIDIA CUDA Toolkit ≥ 12.0 with cuBLAS, cuSOLVER,
  cuRAND

CUDA is *optional*. The package builds, installs, and passes its full
test suite without it.

## What didgpu installs alongside

didgpu depends on a small, stable set of packages — `data.table`,
`Rcpp`/`RcppEigen`, `MASS`, `sandwich`, `jsonlite`. It deliberately does
**not** import `polars` or any of the reference packages it benchmarks
against; those are in `Suggests` and only loaded when you ask for the
`backend = "reference"` parity mode (used for testing).

Go to the next chapter to fit your first model.
