---
title: "Quick start"
---

# Quick start

Five minutes, one panel, one fit, one event-study plot. This chapter
uses the simulator that ships with the package so you can run every
line of code yourself without any extra data.

## Generate a panel

```{r setup}
library(didgpu)

panel <- didgpu_simulate_panel(
  n_units          = 60,
  n_periods        = 12,
  frac_treated     = 0.6,
  min_treat_period = 4,
  max_treat_period = 9,
  tau_profile      = c(0.5, 1.0, 1.2, 1.0, 0.8),
  sigma            = 0.4,
  seed             = 17
)

head(panel)
```

`panel` is a tidy long data frame with columns `unit`, `period`, `Y`,
`D`. About 60% of the units switch from control to treated at some
period between 4 and 9; the true event-study profile (relative to the
switch period) is `c(0.5, 1.0, 1.2, 1.0, 0.8)`.

## Fit the dynamic DiD estimator

The flagship estimator in didgpu is `didgpu()` — the heterogeneity-
robust dynamic DiD of de Chaisemartin & D'Haultfoeuille (2024).

```{r fit}
fit <- didgpu(
  panel,
  outcome    = "Y",
  group      = "unit",
  time       = "period",
  treatment  = "D",
  effects    = 4,
  placebo    = 2,
  backend    = "r",       # try "cuda" once you've checked didgpu_has_cuda_support()
  verbose    = FALSE
)

fit
```

You get:

* **Effects** — a per-event-time matrix with the point estimate, SE,
  CI, and the four sample-size columns the reference reports
  (unweighted N, unweighted switchers, weighted N, weighted switchers).
* **Placebos** — the pre-treatment falsification estimates at horizons
  $-1, -2, \dots$.
* **ATE** — the cumulative incidence-weighted ATE across event-times.
* **Joint tests** — chi-square joint tests over the effects vector
  and over the placebos vector, using the bootstrap empirical
  covariance.

## Add a bootstrap

The point estimate is essentially free; standard errors and CIs come
from a cluster bootstrap. You opt in with `bootstrap_reps`:

```{r fit-boot, eval=FALSE}
fit_boot <- didgpu(
  panel,
  outcome   = "Y", group = "unit", time = "period", treatment = "D",
  effects   = 4,
  placebo   = 2,
  bootstrap_reps = 500,
  backend   = "r",
  seed      = 42,
  verbose   = TRUE
)
```

For a panel this small the bootstrap is fast on CPU. For larger panels,
this is where the GPU pays for itself — see
[Chapter 8 — GPU & Performance](08-gpu-and-performance.qmd).

## Plot the event study

```{r plot, eval=FALSE}
plot(fit)
```

The `plot()` method draws the placebos on negative horizons and the
effects on positive horizons, with the stored CI as error bars and a
horizontal dashed line at zero.

## Pull pieces out with broom-style helpers

```{r tidy-glance, eval=FALSE}
broom::tidy(fit)           # one-row-per-coef data.frame
broom::glance(fit)         # one-row summary (N, n_switchers, joint p)
coef(fit)                  # named numeric of point estimates
vcov(fit)                  # bootstrap empirical covariance
confint(fit, level = 0.95) # CIs at any level (re-fit if level differs)
```

If broom is not loaded, the equivalents `didgpu_tidy()` and
`didgpu_glance()` work the same way without it.

## Save a fitted model

```{r save, eval=FALSE}
saveRDS(fit, "did_fit.rds")
readRDS("did_fit.rds")
```

A didgpu fit is a plain S3 object on top of a list of matrices — it
survives `saveRDS` / `readRDS` round-trips and stays portable across
machines.

## Where to go next

* **[Chapter 3 — Dynamic DID](03-dynamic-did.qmd)** — every option of
  `didgpu()` with worked examples.
* **[Chapter 8 — GPU & Performance](08-gpu-and-performance.qmd)** —
  how to make this 50× faster.
* **[Chapter 10 — Reproducibility & checkpointing](10-reproducibility-checkpointing.qmd)** —
  how to make a 12-hour bootstrap survive a crash.
