---
title: "Sensitivity analysis"
---

# Sensitivity analysis

didgpu ships with three sensitivity-analysis wrappers that take any
fitted didgpu / didgpu_cs / didgpu_fect object and return the same
robust-confidence-set output the underlying methodology papers do.

## HonestDiD (relative magnitudes, smoothness)

```{r honest, eval=FALSE}
hd <- didgpu_honest_did(fit, M_bar = c(0, 0.5, 1, 2))
hd
plot(hd)
```

Wraps [@rambachan2023more]'s HonestDiD package. The "smoothness"
($M$) and "relative magnitudes" ($\bar{M}$) restrictions both work
unchanged; the FLCI variant is also available.

## Rambachan-Roth equivalence (passes-at-delta)

```{r equiv, eval=FALSE}
eq <- didgpu_equivalence(fit, delta = 0.05, alpha = 0.05)
eq
```

The "equivalence-style" test reports `passes_at_delta` — whether the
joint test rejects the null that *every* effect is within $\pm\delta$
of zero.

## Freyaldenhoven-Hansen-Shapiro pre-trend correction

```{r fhs, eval=FALSE}
fhs <- didgpu_freyaldenhoven(fit, ...)
```

Wraps [@freyaldenhoven2019pretrends] for trend-stable confidence sets.

::: {.callout-tip}
A worked walkthrough — taking one fit through all three
sensitivity-analysis methods and reading the output — is in progress.
:::

## References

::: {#refs}
:::
