---
title: "Reproducibility & checkpointing"
---

# Reproducibility & checkpointing

didgpu's checkpointing is the difference between a 12-hour bootstrap
that resumes after a crash and a 12-hour bootstrap that has to start
over.

## Turn it on

```{r checkpoint, eval=FALSE}
fit <- didgpu(
  panel, ...,
  bootstrap_reps = 2000,
  checkpoint_dir = "~/didgpu-runs/2026-05-28-fit",
  seed           = 17
)
```

didgpu writes one small RDS file per bootstrap cell (the per-rep
sufficient statistics, not the panel — usually a few KB each).

## Resume after a crash

Just rerun the **same call** with the **same** `checkpoint_dir` and
`seed`. didgpu reads the existing cells, picks up at the next missing
one, and finishes:

```{r resume, eval=FALSE}
fit <- didgpu(
  panel, ...,
  bootstrap_reps = 2000,
  checkpoint_dir = "~/didgpu-runs/2026-05-28-fit",  # same path
  seed           = 17                                # same seed
)
```

No `tryCatch`, no manual sharding, no "what was I up to?" archaeology.

## How "same seed = same result" works

didgpu's bootstrap uses a deterministic per-cell seed derived from
`(seed, replicate_index)`. So:

* Cells produced before the crash are bit-identical to what the resumed
  run would produce.
* The aggregator sees a complete cell set regardless of which run
  wrote each one.
* Bootstrap point estimate, SE, CI, and joint p-values are
  reproducible across crashes.

## Aggregating already-saved cells

If you want to look at a partial run without re-fitting:

```{r aggregate, eval=FALSE}
partial <- didgpu_aggregate_cells(
  checkpoint_dir = "~/didgpu-runs/2026-05-28-fit"
)
partial
```

Returns the aggregated fit using whatever cells exist on disk. Useful
for sanity-checking convergence while a long run is still going.

::: {.callout-tip}
A worked example — deliberately interrupt a 1000-rep bootstrap, then
resume it, then compare to a fresh run — is in progress.
:::
