---
title: "Bootstrap variants & clustering"
---

# Bootstrap variants & clustering

didgpu supports two bootstrap kinds for `didgpu_cs()` and `didgpu_fect()`:

| Kind | Description | When to use |
|---|---|---|
| `"cluster"` | unit-level resampling, fit refit per replicate | classical, agnostic about the inference. Slow at large B. |
| `"multiplier"` | influence-function multiplier on per-unit IFs | much faster (no refit per replicate). Requires the IF representation, which all estimators in didgpu have. |

For `didgpu()` only the cluster bootstrap is supported (the dynamic
estimator's IFs require additional work that is on the roadmap).

## Specifying a cluster

By default the bootstrap clusters on the unit:

```{r cluster-default, eval=FALSE}
fit <- didgpu(..., bootstrap_reps = 500)
```

If your unit is *not* the natural cluster (panel of student outcomes
where the policy varies at the school level, etc.), pass `cluster`:

```{r cluster-explicit, eval=FALSE}
fit <- didgpu(
  panel, ...,
  cluster        = "school_id",
  bootstrap_reps = 500
)
```

## Resampling at multiple levels

For multi-way clustering (e.g. classroom-within-school), use a
composite cluster id:

```{r multi-way, eval=FALSE}
panel$cluster_composite <- paste(panel$school_id, panel$classroom_id, sep = ":")
fit <- didgpu(panel, ..., cluster = "cluster_composite", bootstrap_reps = 500)
```

didgpu's resampler is a 1-pass random hash of clusters → reps, so it
scales linearly in B.

## Multiplier bootstrap

```{r multiplier, eval=FALSE}
cs <- didgpu_cs(
  panel, ...,
  est_method     = "DR",
  bootstrap_reps = 5000,
  bootstrap_kind = "multiplier",
  seed           = 17
)
```

This is the fast variant. It uses per-unit influence functions
$\psi_i$ and Rademacher / Mammen multipliers $\xi_b$:

$$
\hat\theta^{(b)} - \hat\theta \approx \frac{1}{n} \sum_i \xi_b \, \psi_i
$$

On the GPU, all of this is a couple of GEMVs and a Cholesky per
multiplier — vastly faster than refitting the inner estimator 5000
times.

::: {.callout-tip}
A worked example comparing cluster and multiplier bootstrap SEs on
the same panel (they agree to ~1e-2 at B = 2000) is in progress.
:::
