---
title: "Getting started with didgpu"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with didgpu}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```

`didgpu` is a checkpointed, GPU-capable reimplementation of the
heterogeneity-robust dynamic difference-in-differences estimator of
de Chaisemartin and D'Haultfoeuille (2024), with bit-for-bit numerical
equivalence to the reference `DIDmultiplegtDYN` package across the
features it supports.

This vignette is a hands-on tour:

1. Simulate a panel with a known treatment-effect profile.
2. Estimate event-study effects with placebos and a bootstrap.
3. Plot, summarise, and extract coefficients with standard R methods.
4. Resume an interrupted run from a disk checkpoint.

```{r}
library(didgpu)
```

## 1. Simulate a panel

`didgpu_simulate_panel()` produces a balanced panel with unit and time
fixed effects, a staggered treatment onset (one-period switcher-in for
each treated unit at a random period in `[min_treat_period,
max_treat_period]`), and a known event-time effect profile.

```{r}
panel <- didgpu_simulate_panel(
  n_units   = 100L,
  n_periods = 20L,
  frac_treated = 0.6,
  tau_profile  = c(0.5, 1.0, 1.2, 1.0, 0.8),  # true ATT at k = 0..4
  sigma = 0.4,
  seed  = 17L
)
head(panel)
```

`attr(panel, "truth")` stores the underlying DGP for later validation:
the realised `F_g` per unit and the requested `tau_profile`.

## 2. Pre-flight panel summary

Before committing to a long bootstrap, it is worth running
`didgpu_summarize_panel()` to sanity-check the design:

```{r}
didgpu_summarize_panel(panel,
                       outcome = "Y", group = "unit",
                       time = "period", treatment = "D")
```

This prints baseline cohort sizes, the number of switchers in each
direction, the F_g distribution, and the feasible event-time and
placebo horizons.

## 3. Estimate with a small bootstrap

For interactive exploration, run a small bootstrap (e.g. 20 reps) to
get a quick read on SEs and CIs:

```{r}
fit <- didgpu(
  panel, outcome = "Y", group = "unit", time = "period", treatment = "D",
  effects = 4L, placebo = 2L,
  bootstrap_reps = 20L, seed = 1L,
  backend = "r", verbose = FALSE
)
print(fit)
```

The `Effects` block shows per-event-time ATT estimates with bootstrap
SE, normal-approximation CIs at the stored `ci_level`, and four sample-
size columns (`N`, `Switchers`, `N.w`, `Switchers.w`) that match the
reference `DIDmultiplegtDYN` output exactly. The `Placebos` block is
the corresponding pre-treatment-only difference; jointly the placebos
should be statistically indistinguishable from zero if the
parallel-trends assumption holds. The `ATE` is the cumulative
incidence-weighted average effect.

## 4. Standard R accessors

`didgpu_result` plays nicely with the standard model-extraction
interface:

```{r}
coef(fit)            # named vector of estimates
confint(fit)         # CI matrix at the stored ci_level
diag(vcov(fit))      # bootstrap variances
```

For a quick base-R event-study plot:

```{r, fig.alt = "Base-R event-study plot of the simulated fit."}
plot(fit, main = "Simulated didgpu fit")
```

Pre-treatment placebos appear at negative horizons, post-treatment
effects at positive horizons, with the stored CIs as vertical error
bars and a dashed line at zero for visual reference.

`broom` methods are also provided:

```{r, eval = requireNamespace("broom", quietly = TRUE)}
broom::tidy(fit)
broom::glance(fit)
```

## 5. Long-running runs with disk checkpoints

For real work, the bootstrap often runs for hours. `didgpu()` will
write every completed iteration to disk if you pass
`checkpoint_dir = ...`; a crash at any point keeps everything already
done:

```{r}
cdir <- tempfile("didgpu_demo_")
fit_ck <- didgpu(
  panel, outcome = "Y", group = "unit", time = "period", treatment = "D",
  effects = 4L, placebo = 2L,
  bootstrap_reps = 10L, seed = 1L,
  checkpoint_dir = cdir,
  backend = "r", verbose = FALSE
)
list.files(file.path(cdir, "cells"))
```

If the process dies, resume with `didgpu_resume()`, which reads every
argument back from `meta.json` so you only need to supply the
checkpoint directory and the panel:

```{r}
fit_resumed <- didgpu_resume(cdir, df = panel)
identical(
  as.numeric(fit_ck$results$Effects[, "Estimate"]),
  as.numeric(fit_resumed$results$Effects[, "Estimate"])
)
```

To extend a finished run with additional bootstrap replications without
redoing existing ones:

```{r}
fit_more <- suppressMessages(
  didgpu_bootstrap_more(cdir, df = panel, extra_reps = 5L)
)
nrow(fit_more$results$Effects)  # same number of event-times
fit_more$args$bootstrap_reps    # now 15
```

```{r cleanup, include = FALSE}
unlink(cdir, recursive = TRUE)
```

## 6. Subgroup estimation

`didgpu_by()` runs `didgpu()` separately for each level of a grouping
variable, with each subgroup getting its own checkpoint subdirectory
if `checkpoint_dir` is set:

```{r}
panel$region <- ifelse(panel$unit %% 2L == 0L, "north", "south")
fit_by <- didgpu_by(
  panel, by_var = "region",
  outcome = "Y", group = "unit", time = "period", treatment = "D",
  effects = 3L, bootstrap_reps = 0L,
  backend = "r", verbose = FALSE
)
print(fit_by)
```

## 7. Choosing a backend

```{r}
didgpu_backend_info()
```

For the supported subset, the `"r"` backend (pure R via `data.table`)
is bit-identical to `"reference"` (which delegates to
`DIDmultiplegtDYN`) on point estimates, SEs, and sample-size columns,
but is roughly 14-60x faster depending on panel size. Use
`backend = "reference"` if you want to fall back to the canonical
implementation explicitly.

The `"cuda"` backend is scaffolded but requires an NVIDIA CUDA Toolkit
install with full headers; see `inst/doc/cuda_setup.md` for the
Windows install path. Once `nvcc` is on `PATH` and the package is
reinstalled from source, `didgpu_backend_info()` will report it as
available and `backend = "cuda"` will run the GPU kernels.

## 8. Comparing against the reference

If you have `DIDmultiplegtDYN` installed and want to verify that
didgpu and the reference produce identical output on your panel:

```{r, eval = requireNamespace("DIDmultiplegtDYN", quietly = TRUE)}
report <- didgpu_compare(panel,
                          outcome = "Y", group = "unit",
                          time = "period", treatment = "D",
                          effects = 4L, placebo = 2L)
report$pass
head(report$report)
```

A `pass = TRUE` result with all `max_abs_diff` near machine epsilon
confirms bit-for-bit agreement on point estimates, ATE, and the four
sample-size columns.

## Further reading

- `?didgpu` for the full argument reference.
- `?didgpu_by`, `?didgpu_resume`, `?didgpu_bootstrap_more` for the
  ergonomic helpers.
- `inst/doc/reference_internals.md` (in the installed package) for the
  algorithmic spec.

## References

de Chaisemartin, C. and X. D'Haultfoeuille (2024). "Difference-in-
differences estimators of intertemporal treatment effects."
*The Review of Economics and Statistics*.
[doi:10.1162/rest_a_01414](https://doi.org/10.1162/rest_a_01414).
